Generalists are the most urban-tolerant of birds: a phylogenetically controlled analysis of ecological and life-history traits using a novel continuous urbanization index

### 2018-06-18 ms.Rmd
more changes/edits/polish

# DISCUSSION
The most urban-tolerant species in our analyses, defined based on their urbanization indices, had large niche breadths, reflected in generalism. Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) seldom occurred in urban environments. These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S4), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts]. Further, the non-random distribution of the urbanization index across the phylogeny (Fig. 3 & Table 2) is consistent of reduced phylogenetic diversity in urban areas [@la2018phylogenetic]. Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow a subset of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].

This strong phylogenetic signal contradicts other studies [@evans2011makes], but was unsurprising given our large sample size of 477 species compared to the next highest of 55 species [@evans2011makes]. We also demonstrated the importance of treating species as non-independent, by considering phylogeny in the context of the linear models, highlighting the importance of investigating trait-urbanization relationships in a phylogenetically informed manner. For example, some of the more important traits in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) had a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models, compared to non-phylogenetic models. This was robust relative to phylogenetic uncertainty (Fig. S4). Other traits showed the opposite trend. Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models. This indicates that within clades, species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] were most suited to urban-tolerance. Some have indicated poor evidence that urban-tolerant species have large residual brain size [@kark2007living; @evans2011makes], but we found it was highly correlated with urban-tolerance [Fig. 4; Fig. 5; @moller2009successful; @maklakov2011brains; @moller2015brain].

Although many traits can identify a species' adaptability to urban environments, the published evidence has remained equivocal. Previously, much of the current focus was on discrete classifications of species, based on their responses to urbanization. Contrastingly, we assessed the applicability of traits relative to one another; we did not run separate models or statistical tests for each specific trait [e.g., @kark2007living], allowing comparable effect sizes. Despite this, there appear to be some consistent and generalizable trends within the trait-urbanization complex across the globe. For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful]. Additionally, urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes]. Although we did not find any strong support for our hypotheses on different nesting types, we found a positive association between the degree of sociality and urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living]. While the aforementioned patterns may be generalizable, we also found patterns in our analyses which seem to contrast the global norm. For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [@jones2018birdfeeding] generally involving different species and foods  [compared to European cities for instance; @fuller2008garden].

Our analysis was made possible through integration of broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen]. This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google]. We did not explicitly account for changes throughout the annual cycle in a species' association with urban environments [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor. We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' urbanization index throughout the full-annual cycle [@hostetler2015full; @marra2015call]. This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species, thus allowing prediction of which species face the greatest risk of impending urbanization, specifically in developing countries.

Fully understanding the suite of traits which are associated with urban-tolerant species will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes]. Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role]. We recommend that a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more]. A species-specific approach, which assesses particular resources necessary for specialist species, for example focusing on foraging [@shochat2004linking] and nesting [@fernandez2001habitat] requirements, is likely necessary. To this end, our urbanization index can be used as a potential means to provide quantifiable targets for local restoration projects. For example, a particular restoration project should aim to add species which have lower urbanization indices into their local avifauna, but we note that these indices may differ at different spatial scales [@suarez2018scale].

We provide a novel continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments. Future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance. Our results suggest that generalist species, with high clutch sizes and large residual brain sizes, are best prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns]. In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management. Urban avian biodiversity could be increased by emphasizing the critical importance of intact vegetation [@sushinsky2013should] for specialist species [@devictor2008distribution], and concomitantly improving the suitability of foraging for granivores and insectivores.

>>>>

# DISCUSSION
The most urban-tolerant species in our analyses, defined based on their urbanization indices, had large niche breadths, reflected in generalism. Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) seldom occurred in urban environments. These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S4), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts]. Further, the non-random distribution of the urbanization index across the phylogeny (Fig. 3 & Table 2) is consistent with reduced phylogenetic diversity in urban areas [@la2018phylogenetic]. Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow specific subsets of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].

This strong phylogenetic signal contradicts other studies [@evans2011makes], but was unsurprising given our large sample size of 477 species compared to the next highest of 55 species [@evans2011makes]. We also demonstrated the importance of treating species as non-independent, by considering phylogeny in the context of the linear models, highlighting the importance of investigating trait-urbanization relationships in a phylogenetically informed manner. For example, some of the more important traits in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) had a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models than non-phylogenetic models. Other traits showed the opposite trend. Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models. This indicates that within clades, species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] were most suited to urban-tolerance. Some have indicated poor evidence that urban-tolerant species have large residual brain size [@kark2007living; @evans2011makes], but we found it was highly correlated with urban-tolerance [Fig. 4; Fig. 5; @moller2009successful; @maklakov2011brains; @moller2015brain].

Although many traits can identify a species' adaptability to urban environments, the published evidence has remained equivocal. Previously, much of the current focus was on discrete classifications of species, based on their responses to urbanization [@geschkecompact]. Contrastingly, we assessed the applicability of traits relative to one another; we did not run separate models or statistical tests for each specific trait [e.g., @kark2007living], allowing comparable effect sizes. There appear to be some consistent and generalizable trends within the trait-urbanization complex across the globe. For instance, our identification of the relative importance of annual fecundity was consistent with work from Europe [@croci2008does; @moller2009successful]. Urbanization also appears to select against insectivorous species [@chace2006urban; @evans2011makes]. The degree of sociality was also generally positively associated with urbanization, as found elsewhere [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living]. However, our study also highlighted differences with global norms. For our data, urbanization selected against granivorous species, not previously identified [cf. @chace2006urban; @kark2007living; @evans2011makes]; this could be an artefact of relatively little supplementary feeding in Australian cities [@jones2018birdfeeding] compared to European cities [@fuller2008garden].

The large scale of our analysis was only possible following successful integration of broad-scale empirical datasets (i.e., citizen science data [@silvertown2009new] and remote sensing technology [@gorelick2017google]), which aid our understanding of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen]. We did not explicitly incorporate temporal changes in relation to the life cycle for different species' [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered. In future, especially within migrant-dominated systems (e.g., Europe, North America), temporal changes in association with our urbanization index would add considerable value [@hostetler2015full; @marra2015call]. Australia has only a recent history of rapid urbanization [@coffee2016visualising] and our results may differ for parts of the world with long histories of development, and at different spatial scales [@suarez2018scale]. Yet, there is an opportunity to test our model predictions in urban environments in other parts of the world, potentially allowing prediction of which species face the greatest risk of impending urbanization.

Understanding which traits promote urban-tolerance or intolerance can inform actions to maintain avian biodiversity in urban environments [@evans2011makes]. In particular, future urban development could maintain biodiversity through improved protection of intact native vegetation [@sushinsky2013should] connected by corridors [@litteral2017role], advantaging specialist species detrimentally affected by landscape fragmentation and disturbance [@devictor2008distribution]. Further, increasing insect abundance and diversity in urban environments is potentially important [@baldock2015uk], given collapsing insect populations around the world [@hallmann2017more]. A species-specific approach could help to enhance urban biodiversity, which assesses foraging [@shochat2004linking] or nesting [@fernandez2001habitat] requirements of each species, for example, used to guide protection, mitigation, and restoration of necessary habitats in cities. Our urbanization index could provide quantifiable targets for local restoration projects, where a project aims to reduce the mean urbanization index for their local avifauna.

In conclusion, our novel continuous urbanization index showed relationships between ecological and life history traits with avian tolerance to urban environments. Improving understanding of the temporal nature of these relationships will be important, but we highlight the importance of considering these questions using a continuous measure of urbanization [@evans2011makes] as well as using a phylogenetically informed approach. It is the generalist species [@shochat2006patterns], with high clutch sizes and large residual brain sizes, which are best prepared to forage and survive in the ecologically novel aspects of urban environments.
